The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba

### 2020-12-04 analysis/paper/paper.Rmd
pdf output

---
title: 'The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba'
author:
  - Sheila Rodriguez-Machado:
      email: srodr62@lsu.edu
      institute: [UofO]
      correspondence: true
  - Tomas M. Rodriguez-Cabrera:
      email: tomasmichelrodriguez@gmail.com
      institute: [UofA]
      correspondence: false
institute:
  - UofO: Museum of Natural Sciences, Louisiana State University
  - UofA: Sociedad Cubana de Zoologia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: '../templates/template.docx'
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: '../templates/journal-of-archaeological-science.csl'
abstract: |
  The lack of diagnostic characters to distinguish among species and even genera of gobies (Gobiidae) is still a challenge for science in the 21st century. However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematics of these fish taxa. The fish fry fishery in Eastern Cuba known as “Teti” represents a local fishery and there is no information about the exploitation of this natural resource outside this region. Fishing “Teti” has been a common practice for several years but an exponential increase of tourism in Eastern Cuba represents an additional threat to our biodiversity. Besides, if “Teti” refers to a single species or a species complex is still a knowledge gap in Cuba. The present study aims to identify the Fish Fry "Teti" Fishery in Eastern Cuba using molecular markers. We conducted a barcoding analysis comparing Cuban samples of gobies with samples of the genera *Sicydium* and *Awaous* available in GenBank. Here, we provide evidence that Busck´s Stone-Biting Goby, *Sicydium buscki*, is a component of the fish fry “Teti” fishery in Eastern Cuba. However, whether this fishery includes a single species or more than one species will remain as a mystery until more studies are conducted. Although this study represents the first effort toward and effective management, more thorough investigations and identification of the species comprising the “Teti” fishery are needed. It is imperative to include them in local management plans and to avoid overfishing this resource.  
keywords: |
  amphidromy; Busck's Stone-Biting Goby; COI; DNA barcoding; Gobiidae

@@

#knitr::opts_knit$set(root.dir = '/rrcompendium/')

@@

```{r}

@@

Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure \@ref(fig:mainMap)), eastern Cuba, in December, 2018. Samples were collected using hand nets (1 mm mesh size) 2 m from the riverbank. Specimens were anesthetized using MS 222 (Tricaine Methanesulphonate) and immediately preserved in 80% ethanol. Voucher specimens were deposited in the collection of the Acuario Nacional de Cuba (catalog number: ANC-xxxx).

```{r mainMap, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba"}

@@

```{r gettingAPI}

@@

```{r editingAPI}

@@

```{r alignment}
invisible({capture.output({                ### ALIGNING the sequences        # libraries "msa" and "Biostrings"

@@

```{r, getRF, warning = FALSE}
invisible({capture.output({               ### GETTING the READING FRAME     --- library "systemPipeR"

@@

```{r getBestModel, warning = FALSE}
                                      # FINDING THE BEST NUCLEOTIDE SUSBSTITUTION MODEL

@@

```{r tree, echo = FALSE, include=TRUE, results = 'hide', fig.show='hide', warning = FALSE}

@@

```{r loglikelihood, include = TRUE, results='hide'}
# this result wont be displayed in the final document
loglikelihood <- round(-1*(as.numeric(fit[1])), 2)

@@

```{r editingTree, echo = FALSE}

@@

The final alignment comprised 209 sequences, each one of 639 base pairs. Figure \@ref(fig:SicydiumTree) depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp. available in GenBank. Six clades matching known *Sicydium* species were recovered with strong support. The phylogenetic reconstruction shows that the six samples from Cuba, highlighted in green, nested within *Sicydium buscki* Evermann & Clark, 1906.

```{r SicydiumTree, fig.width = 15, fig.height = 19, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*. Sequences highlited in green represent samples from Cuba. Numbers on branches represent boostrap values higher than 82.")}

@@

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

<div id="refs"></div>

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

>>>>

---
title: 'The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba'
author:
  - Sheila Rodriguez-Machado:
      email: srodr62@lsu.edu
      institute: [UofO]
      correspondence: true
  - Tomas M. Rodriguez-Cabrera:
      email: tomasmichelrodriguez@gmail.com
      institute: [UofA]
      correspondence: false
institute:
  - UofO: Museum of Natural Sciences, Louisiana State University
  - UofA: Sociedad Cubana de Zoologia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    pdf_document:
      toc: FALSE
      fig_caption: yes
      citation_package: biblatex
      latex_engine: xelatex
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: '../templates/journal-of-archaeological-science.csl'
abstract: |
  The lack of diagnostic characters to distinguish among species and even genera of gobies (Gobiidae) is still a challenge for science in the 21st century. However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematics of these fish taxa. The fish fry fishery in Eastern Cuba known as “Teti” represents a local fishery and there is no information about the exploitation of this natural resource outside this region. Fishing “Teti” has been a common practice for several years but an exponential increase of tourism in Eastern Cuba represents an additional threat to our biodiversity. Besides, if “Teti” refers to a single species or a species complex is still a knowledge gap in Cuba. The present study aims to identify the Fish Fry "Teti" Fishery in Eastern Cuba using molecular markers. We conducted a barcoding analysis comparing Cuban samples of gobies with samples of the genera *Sicydium* and *Awaous* available in GenBank. Here, we provide evidence that Busck´s Stone-Biting Goby, *Sicydium buscki*, is a component of the fish fry “Teti” fishery in Eastern Cuba. However, whether this fishery includes a single species or more than one species will remain as a mystery until more studies are conducted. Although this study represents the first effort toward and effective management, more thorough investigations and identification of the species comprising the “Teti” fishery are needed. It is imperative to include them in local management plans and to avoid overfishing this resource.  
keywords: |
  amphidromy; Busck's Stone-Biting Goby; COI; DNA barcoding; Gobiidae

@@

@@

```{r, echo = TRUE}

@@

Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure 1), eastern Cuba, in December, 2018. Samples were collected using hand nets (1 mm mesh size) 2 m from the riverbank. Specimens were anesthetized using MS 222 (Tricaine Methanesulphonate) and immediately preserved in 80% ethanol. Voucher specimens were deposited in the collection of the Acuario Nacional de Cuba (catalog number: ANC-xxxx).

```{r mainMap, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba", echo = TRUE}

@@

```{r gettingAPI, echo = TRUE}

@@

```{r editingAPI, echo = TRUE}

@@

```{r alignment, echo = TRUE}
invisible({capture.output({                ### ALIGNING the sequences        # libraries "msa" and "Biostrings"

@@

```{r, getRF, warning = FALSE, echo = TRUE}
invisible({capture.output({               ### GETTING the READING FRAME     --- library "systemPipeR"

@@

```{r getBestModel, warning = FALSE, echo = TRUE}
                                      # FINDING THE BEST NUCLEOTIDE SUSBSTITUTION MODEL

@@

```{r tree, echo = FALSE, include=TRUE, results = 'hide', fig.show='hide', warning = FALSE, echo = TRUE}

@@

```{r loglikelihood, include = TRUE, results='hide', echo = TRUE}
# this result wont be displayed in the final document
loglikelihood <- round(-1*(as.numeric(fit[1])), 2)

@@

```{r editingTree, echo = FALSE, fig.keep='none', echo = TRUE}

@@

The final alignment comprised 209 sequences, each one of 639 base pairs. Figure 2 depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp. available in GenBank. Six clades matching known *Sicydium* species were recovered with strong support. The phylogenetic reconstruction shows that the six samples from Cuba, highlighted in green, nested within *Sicydium buscki* Evermann & Clark, 1906.

```{r SicydiumTree, fig.width = 15, fig.height = 19, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*. Sequences highlited in green represent samples from Cuba. Numbers on branches represent boostrap values higher than 82."), echo = TRUE}

@@
